---
title: "Effect of stay-at-home order on Covid-19 infectious population growth rate in 5 European countries: France, Germany, UK, Italy, and Poland"
author: "Team 2: Zeng Fung Liew (913802324), Yixing Lu (915501254), Liela Meng (917843295), Apurv Srivastav (918936075)"
date: "March 5th, 2021"
output:
  html_document:
    df_print: paged
    number_sections: yes
---
```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.pos = 'H')
```

# Abstract 



# Introduction


 
# Background 



# Descriptive analysis 
The summary statistics should include at least time, number of cases, number of death, case-mortality rate.

# Inferential analysis 

## Lagged effect of stay-at-home order 

## Effect of policy strictness on infectious population growth rate

# Sensitivity analysis 

## Lagged effect of stay-at-home order 

## Effect of policy strictness on infectious population growth rate


# Causal interpretation 


# Discussion 


# Acknowledgement {-}


# Reference {-}


# Session info {-}

<span style='color:blue'>
Report information of your `R` session for reproducibility. 
</span> 


```{r}
sessionInfo()
```