---
title: "Project3_response"
author: "YL"
date: "2/25/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(dplyr)
library(zoo)
```

```{r}
# load data 
covid <- read_csv("https://covid19.who.int/WHO-COVID-19-global-data.csv")
covid[which(covid$Country == "The United Kingdom"), "Country"] = "UK"
write_csv(covid,file="WHO-Covid-19-backup.csv")

# filter out countries in "Other" WHO regions
covid <- covid %>% 
  filter(WHO_region != "Other") %>% 
  mutate(WHO_region = fct_recode(WHO_region,
                                 "Eastern Mediterranean"="EMRO",
                                 "Europe" = "EURO",
                                 "Africa" = "AFRO",
                                 "Western Pacific" = "WPRO",
                                 "Americas"="AMRO",
                                 "South-East Asia" = "SEARO"))

# stay-at-home dataset
home = read_csv("stay-at-home-covid.csv")             # change filename/directory if necessary
home$stay_home_requirements = as.factor(home$stay_home_requirements)
home[which(home$Entity == "United Kingdom"), "Entity"] = "UK"

# inner join data
covid = inner_join(covid, home, 
                     by = c("Date_reported" = "Date",
                            "Country" = "Entity"),
                     keep = FALSE)
covid = covid[, c("Date_reported", 
                  "Country_code", 
                  "Country", 
                  "WHO_region", 
                  "New_cases", 
                  "Cumulative_cases", 
                  "New_deaths", 
                  "Cumulative_deaths", 
                  "stay_home_requirements")]

# Target period of analysis: 2020-10-01 to 2020-11-30
# Target country: Germany, UK, France, Italy, Poland
covid_window <- covid %>% 
  filter(Country %in% c("Germany", "UK", "France", "Italy", "Poland") & 
           Date_reported>"2020-08-15" & 
           Date_reported<"2020-12-07")

# virus incubation period d = 7
# I_(t-d-1) = y_(t-1) - y_(t-d-1), where y is cumulative cases
# delta_I_t = [y_t - y_(t-d)] - [y_(t-1)-y_(t-d-1)] = new_cases_t - new_cases_(t-d)
covid_model.d7 <- covid_window %>% 
  group_by(Country) %>% 
  mutate(New_infectious = New_cases - lag(New_cases,7),
         lag_infectious = Cumulative_cases - lag(Cumulative_cases,7))

# change in stay home restriction
covid_model.d7 <- covid_model.d7 %>%
  group_by(Country) %>%
  mutate(stay_home_diff_07 = as.factor(as.numeric(stay_home_requirements) - lag(as.numeric(stay_home_requirements), 7)),
         stay_home_diff_05 = as.factor(as.numeric(stay_home_requirements) - lag(as.numeric(stay_home_requirements), 5)),
         stay_home_diff_03 = as.factor(as.numeric(stay_home_requirements) - lag(as.numeric(stay_home_requirements), 3))
  )

# lagged change in stay home restriction
covid_model.d7 <- covid_model.d7 %>%
  group_by(Country) %>%
  mutate(lag_stay_home_diff_07 = as.factor(lag(as.numeric(stay_home_requirements), 30) -
                                         lag(as.numeric(stay_home_requirements), 37)),
         lag_stay_home_diff_05 = as.factor(lag(as.numeric(stay_home_requirements), 30) -
                                         lag(as.numeric(stay_home_requirements), 35)),
         lag_stay_home_diff_03 = as.factor(lag(as.numeric(stay_home_requirements), 30) -
                                         lag(as.numeric(stay_home_requirements), 33))
  )

# calculating 3, 5, 7-day moving average of new_infectious and cum_infectious
covid_model.d7 <- covid_model.d7 %>% 
  group_by(Country) %>% 
  mutate(New_infec_07ma = zoo:: rollmean(New_infectious,k=7,fill=NA),
         lag_infec_07ma = zoo:: rollmean(lag_infectious,k=7,fill=NA),
         New_infec_05ma = zoo:: rollmean(New_infectious,k=5,fill=NA),
         lag_infec_05ma = zoo:: rollmean(lag_infectious,k=5,fill=NA),
         New_infec_03ma = zoo:: rollmean(New_infectious,k=3,fill=NA),
         lag_infec_03ma = zoo:: rollmean(lag_infectious,k=3,fill=NA)) %>%
  dplyr::ungroup()

covid_model.d7 <- covid_model.d7 %>% 
  filter(covid_model.d7$Date_reported>"2020-09-30" & 
           covid_model.d7$Date_reported<"2020-12-01")
range(covid_model.d7$Date_reported) # 2020-10-01 to 2020-11-30

# Calculate response variable: delta_infectious
covid_model.d7 <- covid_model.d7 %>% 
  mutate(delta_infectious = 100*New_infectious/lag_infectious, 
         delta_infec_07ma = 100*New_infec_07ma/lag_infec_07ma, 
         delta_infec_05ma = 100*New_infec_05ma/lag_infec_05ma, 
         delta_infec_03ma = 100*New_infec_03ma/lag_infec_03ma)

# plot of delta_infectious vs. date
plot(covid_model.d7$Date_reported,covid_model.d7$delta_infec_05ma, 
     type="p",cex=0.3,
     main = expression(paste("%", Delta, y[ct], " by Date")),
     xlab = "Date",
     ylab = expression(paste("%", Delta, y[ct])))

write.csv(covid_model.d7,"who_new_response.csv")
```

```{r}
ggplot(data = covid_model.d7, aes(x = Date_reported, y = delta_infec_05ma, col = Country)) +
  geom_point(alpha = I(0.5)) + 
  geom_smooth(method = "loess", formula = y ~ x,  span = 0.5) +
  labs(x = "Date", y = expression(paste("%", Delta, y[ct])), col = "Country") +
  ggtitle(expression(paste("%", Delta, y[ct], " by Date")))
```

Notes:

* Generally trends upwards until October 15 before trending downwards

Bar charts
```{r}
ggplot(
  data = covid_model.d7,
  mapping = aes(
    x = Date_reported,
    y = New_cases,
    fill = stay_home_requirements
  )
) +
  geom_bar(
    stat = "identity"
  ) +
  labs(
    title = "COVID-19 daily cases and stay-at-home requirement in Oct-Nov 2020",
    x = "Date",
    y = "Country",
    fill = "Restriction"
  ) +
  facet_grid(
    rows = vars(Country),
    scales = "free"
  ) +
  theme(
    legend.position = "bottom"
  )
```

Notes:

* All countries increased restriction at around the mid-week of October.

# Models

**Model 1:** Correlation between infectious rate and change in restriction level.
$$
\Delta y_{ct} = \beta_0 + \beta_1 x_{1,t} + \beta_2 x_{2,-1,ct} + \beta_3 x_{2,1,ct} + \beta4 x_{2,2,ct} + \epsilon_{ct}
$$

Note:

* If we compare the following models, the `stay_home_diff_07` model turns out to be the most significant. This is shown in the p-values and $R^2$ values. <br>
* Large increase in infectious rate is highly correlated with increase in stay home restrictions.

```{r}
library(plm)
model1.07 = plm(delta_infec_07ma ~ Date_reported + stay_home_diff_07, data = covid_model.d7,
                index = "Country", model = "within")
summary(model1.07)
```

```{r}
model1.05 = plm(delta_infec_05ma ~ Date_reported + stay_home_diff_05, data = covid_model.d7,
                index = "Country", model = "within")
summary(model1.05)
```

```{r}
model1.03 = plm(delta_infec_03ma ~ Date_reported + stay_home_diff_03, data = covid_model.d7,
                index = "Country", model = "within")
summary(model1.03)
```


**Model 2:** 
Correlation between infectious rate and change in restriction level a month prior.
$$
\Delta y_{ct} = \beta_0 + \beta_1 x_{1,t} + \beta_2 x_{2,1,ct} + \beta_3 x_{2,2,ct} + \epsilon_{ct}
$$

Note: 

* Increasing stay home restrictions DOES work, but there is a time lag of about a month. <br>
* A significant increase in stay home restrictions (up by 2 levels) is necessary for effects to be significant.

```{r}
model1.lag07 = plm(delta_infec_07ma ~ Date_reported + lag_stay_home_diff_07, data = covid_model.d7,
                index = "Country", model = "within")
summary(model1.lag07)
```
